Interpolants, Error Bounds, and Mathematical Software for Modeling and Predicting Variability in Computer Systems

### 2020-08-29 3-Dissertation/Paper/1-apps.tex
Final thesis document completed.

This thesis is comprised of work from multiple conference and journal
papers. Chapters \ref{ch:apps} and \ref{ch:algs} share the broader
context for this work and describe relevant existing algorithms for
approximation, these chapters derive from multiple published papers
\cite{lux2018predictive,lux2020effective,lux2020interpolation}. Chapter
\ref{ch:naive} presents a first attempt at modeling variability and
appears in \cite{lux2018predictive}. Chapter \ref{ch:boxes} appears in
\cite{lux2018novel} and presents multiple novel approximation
techniques. Chapter \ref{ch:strong} applies the novel approximation
techniques to distribution prediction and appears in
\cite{lux2018nonparametric}. Chapter \ref{ch:error} presents a novel
error bound and is to appear in \cite{lux2020interpolation}. Finally
Chapter \ref{ch:splines} proposes a mathematical software package for
monotone quintic spline interpolation, was partially presented in
\cite{lux2020algorithm}, and is to be submitted in full to ACM
Transactions on Mathematical Software (TOMS).

>>>>

This thesis is comprised of work from multiple conference and journal
papers. Chapters \ref{ch:apps} and \ref{ch:algs} share the broader
context for this work and describe relevant existing algorithms for
approximation, these chapters derive from multiple published papers
\cite{lux2018predictive,lux2020effective,lux2020interpolation}. Chapter
\ref{ch:naive} presents a first attempt at modeling variability and
appears in \cite{lux2018predictive}. Chapter \ref{ch:boxes} appears in
\cite{lux2018novel} and presents multiple novel approximation
techniques. Chapter \ref{ch:strong} applies the novel approximation
techniques to distribution prediction and appears in
\cite{lux2018nonparametric}. Chapter \ref{ch:error} presents a novel
error bound and is to appear in \cite{lux2020interpolation}.  Chapter
\ref{ch:splines} proposes a mathematical software package for monotone
quintic spline interpolation, was partially presented in
\cite{lux2020algorithm}, and is to be submitted in full to ACM
Transactions on Mathematical Software (TOMS). Finally Chapter
\ref{ch:concs} concludes, presenting the major takeaways of this work
as well as promising future research directions.

### 2020-08-29 3-Dissertation/Paper/8-conclusions.tex
Final thesis document completed.

This body of work reveals many interesting findings related to modeling variability in computer systems and approximation in general. While generic ``off the shelf'' techniques from machine learning are immediately capable of being applied to the prediction of variability summary statistics, a little extra effort can lend significant improvements in accuracy and precision. The major takeaways from this work are threefold:

@@

Lastly, the improvements in distribution approximation provided by the \texttt{MQSI} software package are not immediately acquired for any data. There is work to be done investigating the best mixture of sampling and smoothing methods to generate a reasonable set of empirical distribution function (EDF) points that can be interpolated by a monotone quintic spline. While initial experiments offered the selection of uniformly distribution EDF points, this may not be optimal and the question of how many points is best remains open. Traditionally Gaussian smoothing or some other method may be used to translate an EDF into a good distribution approximation, however these methods require the (often arbitrary) selection of multiple parameters. Ideally, one would refine the EDF to a small set of points with high confidence that would afford the best possible estimates of first and second derivatives by the quadratic facet model in \texttt{MQSI}.

>>>>

This body of work reveals many interesting findings related to modeling variability in computer systems and approximation in general. While generic ``off the shelf'' techniques from machine learning are immediately capable of being applied to the prediction of variability summary statistics, a little extra effort can lend significant improvements in the accuracy and descriptiveness of the predictions being made. For example, interpolants like Delaunay yield lower error predictions with less data (improved accuracy), while also trivially extending to the prediction of full distributions (improved precision). The major takeaways from this work are threefold:

@@

Lastly, the potential improvements in distribution approximation provided by the \texttt{MQSI} software package are significant. The ability to capture second derivative information in a piecewise monotone spline interpolant can greatly reduce the amount of data required to estimate a distribution from a sample. Extensions of this work should consider statistical importance sampling, smoothing, and other methods to provide robust estimates of points on the distribution functions that can be used in conjunction with \texttt{MQSI} to provide the best possible distribution estimates.

### 2020-08-29 3-Dissertation/Paper/thesis.tex
Final thesis document completed.

% The dedication and acknowledgement pages are optional. Comment them out to remove them.
%% \dedication{...}
%% \acknowledge{...}

@@

@@

% Add your appendices here. You must leave the appendices enclosed in the appendices environment in order for the table of contents to be correct.
\begin{appendices}
  \chapter{Error Bound Appendices} \label{app:error}

\begin{normalsize}
\textbf{Statistical Terminology.} A random variable $X$ is precisely
defined by its \textit{cumulative distribution function} (CDF) $F_X$
and the derivative of the CDF, the \textit{probability density
  function} (PDF) $f_X.$ For any possible value $x$ of $X$, the
\textit{percentile} of $x$ is $100$ $F_X(x)$, the percentage of values
drawn from $X$ that would be less than or equal to $x$ as the number
of samples tends towards infinity. The \textit{quartiles} of $X$ are
its $25$-th, $50$-th (median), and $75$-th percentiles. The absolute
difference between the median and an adjacent quartile is an
\textit{interquartile range}. Given an independent and identically
distributed sample from $X$ and presuming that $X$ has finite mean and
variance, a \textit{confidence interval} can be drawn about any
percentile estimated from the sample. A \textit{confidence interval}
describes the probability that a value lies within an interval. The
\textit{null hypothesis} is a statement (derived from some test
statistic) that the expected value of the observed statistic is equal
to an assumed population statistic. The $p$ value is the probability
of observing a given statistic if the \textit{null hypothesis} is
true. For a more detailed introduction to statistics and 
related terminology, see the work of Navidi \cite{navidi_2015}.

\vspace{4mm}
\noindent \textbf{Raw Numerical Results.} The tables that follow show the
precise experimental results for all data sets presented in Section
\ref{sec:data}. The tests were all run serially on an otherwise idle
machine with a CentOS 6.10 operating system and an Intel i7-3770 CPU
operating at 3.4 GHz. The detailed performance results in the tables
that follow are very much dependent on the problem and the algorithm
implementation (e.g., some codes are TOMS software, some industry
distributions, and others are from conference paper venues). Different
typeface is used to show best performers, however not much
significance should be attached to ranking algorithms based on small
time (millisecond) differences. The results serve as a demonstration
of conceptual validity.

\end{normalsize}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth,]{Figures/NA/scatter_forestfires.pdf}
  \hspace{.6cm}
  \includegraphics[width=0.45\textwidth,]{Figures/NA/scatter_parkinsons.pdf}
  \vspace{.6cm}

  \includegraphics[width=0.45\textwidth,]{Figures/NA/scatter_weather.pdf}
  \hspace{.6cm}
  \includegraphics[width=0.45\textwidth,]{Figures/NA/scatter_creditcard.pdf}
  \vspace{.3cm}

  \caption{Scatter plots for predicted versus actual values for the
    top three models on each of the four real valued approximation
    problems. Top left is forest fire data, top right is Parkinson's
    data, bottom left is rainfall data, and bottom right is credit
    card transaction data. The blue circles correspond to the best
    model, green diamonds to the second best, and red crosses to the
    third best model for each data set. There are a large of number of
    0-valued entries in the forest fire and rainfall data sets that
    are not included in the visuals making the true ranking of the
    models appear to disagree with the observed outcomes.}
  \label{fig:scatter_plots}
\end{figure}

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}       & \multicolumn{2}{c}{$25^{th}$} & \multicolumn{2}{c}{$50^{th}$} & \multicolumn{2}{c}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
    \hline
    MARS      & $0$&$00984$                   & $3$&$11$                      & $7$&$01$                      & $\mathit{11}$&$\mathit{7}$    & $1090$&$0$\\
    SVR       & $0$&$0118$                    & $\mathbf{0}$&$\mathbf{615}$   & $\mathbf{0}$&$\mathbf{931}$   & $\mathbf{5}$&$\mathbf{89}$    & $1090$&$0$\\
    MLP       & $0$&$0426$                    & $2$&$63$                      & $5$&$27$                      & $14$&$0$                      & $1090$&$0$\\
    Delaunay  & $\mathbf{0}$&$\mathbf{0}$     & $1$&$98$                      & $5$&$37$                      & $13$&$1$                      & $\mathit{1080}$&$\mathit{0}$\\
    ShepMod   & $\mathbf{0}$&$\mathbf{0}$     & $1$&$93$                      & $6$&$27$                      & $16$&$0$                      & $1090$&$0$\\
    LSHEP     & $0$&$04$                      & $2$&$17$                      & $8$&$87$                      & $19$&$1$                      & $\mathbf{1070}$&$\mathbf{0}$\\
    Voronoi   & $\mathit{0}$&$\mathit{00982}$ & $3$&$65$                      & $7$&$56$                      & $15$&$6$                      & $1090$&$0$\\
    BoxSpline & $\mathbf{0}$&$\mathbf{0}$     & $\mathit{1}$&$\mathit{27}$    & $\mathit{4}$&$\mathit{61}$    & $12$&$6$                      & $1090$&$0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-forest-fire}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \vspace{.2cm}
  \label{table:error-forest-fire}

  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best}   &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS      & \quad$7$&$3$                   &  & \quad\quad$29$&$1$                      & \quad$0$&$00137$                   & \quad\quad\quad$0$&$0686$\\
    SVR       & $\mathbf{78}$&$\mathbf{0}$     &  & $\mathbf{0}$&$\mathbf{00584}$           & $\mathit{0}$&$\mathit{0000620}$    & $\mathit{0}$&$\mathit{00310}$\\
    MLP       & $0$&$0$                        &  & $32$&$8$                                & $0$&$000871$                       & $0$&$0436$\\
    Delaunay  & $0$&$2$                        &  & $0$&$0151$                              & $0$&$0234$                         & $1$&$18$\\
    ShepMod   & $2$&$0$                        &  & $\mathit{0}$&$\mathit{00634}$           & $0$&$0000644$                      & $0$&$00322$\\
    LSHEP     & $5$&$1$                        &  & $0$&$0275$                              & $\mathbf{0}$&$\mathbf{0000463}$    & $\mathbf{0}$&$\mathbf{00231}$\\
    Voronoi   & $0$&$0$                        &  & $0$&$0152$                              & $0$&$000396$                       & $0$&$0198$\\
    BoxSpline & $\mathit{9}$&$\mathit{7}$      &  & $0$&$00724$                             & $0$&$0000978$                      & $0$&$00489$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating forest fire data in Table
    \ref{table:error-forest-fire}. On the right columns are median fit
    time of 454 points, median time for one approximation, and median
    time approximating 50 points.}
  \label{table:best-forest-fire}
\end{table}
\vspace{-1cm}

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}                   & \multicolumn{2}{c}{$25^{th}$} & \multicolumn{2}{c}{$50^{th}$} & \multicolumn{2}{c}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
    \hline
    MARS      & $0$&$00948$                               & $9$&$98$                      & $20$&$4$                      & $32$&$9$                      & $863$&$0$\\
    SVR       & $0$&$00233$                               & $3$&$15$                      & $7$&$21$                      & $11$&$3$                      & $\mathbf{28}$&$\mathbf{6}$\\
    MLP       & $0$&$0000239$                             & $\mathit{0}$&$\mathit{533}$   & $\mathit{1}$&$\mathit{25}$    & $\mathbf{2}$&$\mathbf{84}$    & $39$&$3$\\
    Delaunay  & $\mathit{3}$&$\mathit{72 \times 10^{-12}}$ & $1$&$2$                       & $3$&$5$                       & $7$&$67$                      & $30$&$7$\\
    ShepMod   & $\mathbf{0}$&$\mathbf{0}$                 & $\mathbf{0}$&$\mathbf{255}$   & $\mathbf{0}$&$\mathbf{908}$   & $\mathit{3}$&$\mathit{43}$    & $34$&$5$\\
    LSHEP     & $0$&$00254$                               & $2$&$93$                      & $7$&$16$                      & $13$&$1$                      & $\mathit{29}$&$\mathit{0}$\\
    Voronoi   & $\mathbf{0}$&$\mathbf{0}$                 & $1$&$29$                      & $3$&$52$                      & $6$&$87$                      & $30$&$1$\\
    BoxSpline & $0$&$006$                                 & $4$&$3$                       & $8$&$91$                      & $15$&$1$                      & $45$&$3$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-parkinsons}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \vspace{.2cm}
  \label{table:error-parkinsons}

  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS      & \quad\,$0$&$0$               &  & \quad\quad\quad$37$&$9$                 & \quad\,\,$0$&$00253$               & \quad\quad\quad\,\,$1$&$48$\\
    SVR       & $0$&$1$                      &  & $\mathbf{0}$&$\mathbf{859}$             & $\mathit{0}$&$\mathit{000181}$     & $\mathit{0}$&$\mathit{106}$\\
    MLP       & $\mathit{32}$&$\mathit{0}$   &  & $348$&$0$                               & $0$&$00111$                        & $0$&$653$\\
    Delaunay  & $0$&$0$                      &  & $2$&$47$                                & $1$&$22$                           & $720$&$0$\\
    ShepMod   & $\mathbf{66}$&$\mathbf{4}$   &  & $\mathit{1}$&$\mathit{13}$              & $0$&$000182$                       & $0$&$107$\\
    LSHEP     & $0$&$0$                      &  & $2$&$39$                                & $\mathbf{0}$&$\mathbf{000144}$     & $\mathbf{0}$&$\mathbf{0845}$\\
    Voronoi   & $1$&$6$                      &  & $2$&$77$                                & $0$&$0274$                         & $16$&$1$\\
    BoxSpline & $0$&$0$                      &  & $1$&$26$                                & $0$&$000643$                       & $0$&$377$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating Parkinson's data in Table
    \ref{table:error-parkinsons}. On the right columns are median fit
    time of 5288 points, median time for one approximation, and median
    time approximating 587 points.}
  \vspace{-1cm}
  \label{table:best-parkinsons}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}                    & \multicolumn{2}{c}{$25^{th}$}              & \multicolumn{2}{c}{$50^{th}$}  & \multicolumn{2}{c}{$75^{th}$}  & \multicolumn{2}{c}{Max}\\
    \hline
    MARS      & $\mathit{6}$&$\mathit{45 \times 10^{-15}}$ & $\mathbf{2}$&$\mathbf{70 \times 10^{-14}}$ & $1$&$66$                       & $1$&$96$                       & $\mathit{53}$&$\mathit{3}$\\
    SVR       & $0$&$0000915$                              & $0$&$0833$                                 & $0$&$263$                      & $\mathit{1}$&$\mathit{13}$     & $109$&$0$\\
    MLP       & $0$&$0000689$                              & $0$&$0337$                                 & $\mathbf{0}$&$\mathbf{0795}$   & $\mathbf{0}$&$\mathbf{264}$    & $\mathbf{5}$&$\mathbf{31}$\\
    Delaunay  & $\mathbf{0}$&$\mathbf{0}$                  & $0$&$187$                                  & $0$&$919$                      & $2$&$72$                       & $56$&$3$\\
    ShepMod   & $\mathbf{0}$&$\mathbf{0}$                  & $0$&$0957$                                 & $0$&$685$                      & $2$&$9$                        & $73$&$2$\\
    LSHEP     & $\mathbf{0}$&$\mathbf{0}$                  & $\mathit{0}$&$\mathit{0153}$               & $\mathit{0}$&$\mathit{106}$    & $1$&$17$                       & $113$&$0$\\
    Voronoi   & $\mathbf{0}$&$\mathbf{0}$                  & $0$&$43$                                   & $1$&$28$                       & $2$&$94$                       & $83$&$8$\\
    BoxSpline & $\mathbf{0}$&$\mathbf{0}$                  & $0$&$342$                                  & $1$&$59$                       & $5$&$53$                       & $119$&$0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-weather}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum value of each column is boldface, while the second lowest
    is italicized. All values are rounded to three significant
    digits.}
  \vspace{.2cm}
  \label{table:error-weather}

  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)}    & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS      & \,\,\,\,$10$&$7$             &  & \quad\quad\quad$\mathit{0}$&$\mathit{151}$ & \quad$0$&$000117$                  & \quad\quad\quad\,\,$0$&$0304$\\
    SVR       & $0$&$0$                      &  & $\mathbf{0}$&$\mathbf{133}$                & $\mathbf{0}$&$\mathbf{0000947}$    & $\mathbf{0}$&$\mathbf{0246}$\\
    MLP       & $\mathbf{60}$&$\mathbf{9}$   &  & $169$&$0$                                  & $0$&$00137$                        & $0$&$356$\\
    Delaunay  & $0$&$1$                      &  & $0$&$664$                                  & $0$&$886$                          & $230$&$0$\\
    ShepMod   & $3$&$5$                      &  & $0$&$265$                                  & $0$&$000128$                       & $0$&$0333$\\
    LSHEP     & $\mathit{28}$&$\mathit{4}$   &  & $0$&$874$                                  & $\mathit{0}$&$\mathit{0000975}$    & $\mathit{0}$&$\mathit{0254}$\\
    Voronoi   & $0$&$2$                      &  & $0$&$675$                                  & $0$&$0270$                         & $7$&$01$\\
    BoxSpline & $4$&$1$                      &  & $0$&$330$                                  & $0$&$000406$                       & $0$&$106$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{Left table shows how often each algorithm had the lowest
    absolute error approximating Sydney rainfall data in Table
    \ref{table:error-weather}. On the right columns are median fit
    time of 2349 points, median time for one approximation, and median
    time approximating 260 points.}
  \vspace{-1cm}
  \label{table:best-weather}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}                    & \multicolumn{2}{c}{$25^{th}$} & \multicolumn{2}{c}{$50^{th}$} & \multicolumn{2}{c}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
    \hline
    MARS      & $5$&$36$                                   & $3610$&$0$                    & $4580$&$0$                    & $5450$&$0$                    & $13400$&$0$\\
    SVR       & $0$&$00706$                                & $8$&$14$                      & $\mathit{13}$&$\mathit{0}$    & $41$&$6$                      & $7690$&$0$\\
    MLP       & $0$&$00151$                                & $\mathbf{1}$&$\mathbf{6}$     & $\mathbf{3}$&$\mathbf{86}$    & $\mathbf{8}$&$\mathbf{34}$    & $\mathbf{604}$&$\mathbf{0}$\\
    Delaunay  & $\mathbf{0}$&$\mathbf{0}$                  & $\mathit{2}$&$\mathit{69}$    & $13$&$8$                      & $\mathit{35}$&$\mathit{0}$    & $\mathit{4840}$&$\mathit{0}$\\
    ShepMod   & $\mathbf{0}$&$\mathbf{0}$                  & $4$&$21$                      & $17$&$3$                      & $51$&$6$                      & $6510$&$0$\\
    LSHEP     & $1$&$27$                                   & $199$&$0$                     & $260$&$0$                     & $343$&$0$                     & $6530$&$0$\\
    Voronoi   & $\mathit{2}$&$\mathit{89 \times 10^{-10}}$ & $14$&$7$                      & $32$&$8$                      & $52$&$9$                      & $4860$&$0$\\
    BoxSpline & $\mathbf{0}$&$\mathbf{0}$                  & $12$&$4$                      & $35$&$0$                      & $90$&$1$                      & $7690$&$0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-credit-card}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum value of each column is boldface, while the second lowest
    is italicized. All values are rounded to three significant
    digits.}
  \vspace{.2cm}
  \label{table:error-credit-card}

  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS      & \quad\,$0$&$0$               &  & \quad\quad\quad$22$&$0$                 & \quad$0$&$00148$                   & \quad\quad\quad\,\,$0$&$820$\\
    SVR       & $0$&$0$                      &  & $\mathbf{1}$&$\mathbf{01}$              & $\mathit{0}$&$\mathit{000210}$     & $\mathit{0}$&$\mathit{117}$\\
    MLP       & $\mathbf{79}$&$\mathbf{5}$   &  & $290$&$0$                               & $0$&$000714$                       & $0$&$397$\\
    Delaunay  & $\mathit{20}$&$\mathit{5}$   &  & $3$&$12$                                & $3$&$71$                           & $2070$&$0$\\
    ShepMod   & $0$&$1$                      &  & $\mathit{1}$&$\mathit{4}5$              & $0$&$000220$                       & $0$&$122$\\
    LSHEP     & $0$&$0$                      &  & $3$&$47$                                & $\mathbf{0}$&$\mathbf{000176}$     & $\mathbf{0}$&$\mathbf{0981}$\\
    Voronoi   & $0$&$0$                      &  & $3$&$32$                                & $0$&$0950$                         & $52$&$8$\\
    BoxSpline & $0$&$0$                      &  & $1$&$66$                                & $0$&$000956$                       & $0$&$532$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating credit card transaction data
    in Table \ref{table:error-credit-card}. On the right columns are
    median fit time of 5006 points, median time for one approximation,
    and median time approximating 556 points.}
  \vspace{-1cm}
  \label{table:best-credit-card}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}      & \multicolumn{2}{c}{$25^{th}$} & \multicolumn{2}{c}{$50^{th}$} & \multicolumn{2}{c}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
    \hline
    Delaunay  & $\mathbf{0}$&$\mathbf{0287}$ & $\mathbf{0}$&$\mathbf{0914}$  & $\mathbf{0}$&$\mathbf{158}$   & $\mathit{0}$&$\mathit{308}$   & $\mathit{0}$&$\mathit{897}$\\
    ShepMod   & $0$&$0307$                   & $\mathit{0}$&$\mathit{0983}$  & $\mathit{0}$&$\mathit{164}$   & $0$&$335$                     & $1$&$0$\\
    Voronoi   & $\mathit{0}$&$\mathit{0303}$ & $0$&$1$                       & $0$&$165$                     & $\mathbf{0}$&$\mathbf{274}$   & $\mathbf{0}$&$\mathbf{762}$\\
    BoxSpline & $0$&$0679$                   & $0$&$429$                     & $0$&$571$                     & $0$&$752$                     & $1$&$0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-throughput}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum KS statistics respectively between
    truth and guess for models predicting the distribution of I/O
    throughput that will be observed at previously unseen system
    configurations. The minimum value of each column is boldface,
    while the second lowest is italicized. All values are rounded to
    three significant digits.}
  \vspace{.2cm}
  \label{table:error-throughput}

  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best}   &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    Delaunay  & \,\,$\mathbf{62}$&$\mathbf{0}$ &  & \quad\quad\quad$0$&$344$                & \quad$0$&$00984$                   & \quad\quad\,\,$2$&$71$\\
    ShepMod   & $0$&$0$                        &  & $\mathbf{0}$&$\mathbf{0884}$            & $\mathbf{0}$&$\mathbf{000145}$     & $\mathbf{0}$&$\mathbf{0436}$\\
    Voronoi   & $\mathit{38}$&$\mathit{0}$     &  & $0$&$360$                               & $0$&$00253$                        & $0$&$762$\\
    BoxSpline & $0$&$0$                        &  & $\mathit{0}$&$\mathit{0972}$            & $\mathit{0}$&$\mathit{000210}$     & $\mathit{0}$&$\mathit{0633}$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest KS statistic on the I/O throughput distribution data in
    Table \ref{table:error-throughput}. On the right columns are
    median fit time of 2715 points, median time for one approximation,
    and median time approximating 301 points.}
  \label{table:best-throughput}
\end{table}

\vspace*{\fill}

>>>>

% The dedication and acknowledgement pages are optional. Comment them out to remove them.
\dedication{
For Alex Faustie Grant, my dream teammate.
}
\acknowledge{
Finishing a Ph.D. is hard work. It's not always fun. And sometimes it's stressful. But the people I've met along the way, the things I've learned, and the people from all paths of life that have supported me throughout this journey made all the difference. My gratitudes are endless, but I'll do my best to enumerate some here.
{\vspace{2mm} \\}
First and foremost I want to thank my advisor Dr. Watson for his enduring push for deep scientific rigor and always reminding me of my most important job, to ``think!'' While reading that without context might fall short, Dr. Watson's incredible effort, earnest care, and impressive repertoire of knowledge afforded me many great lessons throughout my graduate tenure that I value dearly. His genuine feedback has helped me become a better researcher and he's someone I won't hesitate to consult in the future.
{\vspace{2mm} \\}
I want to thank my coadvisor Dr. Hong for deepening my knowledge of statistics and introducing me to countless relevant and interesting lines of research. His support and direction drove the statistical significance of my work. In addition I want to thank all of my committee members for their care and contributions, Dr. Cameron for asking important practical questions and encouraging a bigger picture cohesion throughout my work, Dr. Huang and Dr. Wang for ensuring theoretical groundedness and relevance to modern research as well as being great teachers, and Dr. Cao for motivating many connections between my work and other areas of applied computational science.
{\vspace{-1mm} \\}
I was uniquely lucky in that the entirety of my work at Virginia Tech was supported by the VarSys project. Those team members, both students and faculty, guided my growth as a researcher and my research direction as a whole. I thank them all for the many conversations about fascinating and diverse topics in computer science, as well as constant feedback and collaboration that broadened my interests and drove me to produce higher quality results.
{\vspace{2mm} \\}
Where would I be without my closest friends through graduate school. I sincerely thank all of Grad Council, those who helped found the group and see it to its great success as a leading computer science graduate council in the nation. I am especially grateful for my lab mates Tyler Chang and Ayaan Kazerouni for the lunches around Blacksburg, encouraging conversations, and general support that made graduate life fun and memorable in ways that it would never have been otherwise.
{\vspace{2mm} \\}
Another thanks goes to my family for their continued love and support, particularly my parents Karen and John Lux, brothers Dylan and Austen, sisters Amy, and Arial, and grandmothers Gooma and Mor Mor. They have made all the difference and I am unbelievably grateful for them all.
{\vspace{2mm} \\}
Lastly I would like to thank my fianc\'e Alex Grant for her enduring love and support over the last four years. I couldn't imagine getting this far without our fun trips to Kentucky, summers together, and Skype dinner and movie dates that made the distance between us from Blacksburg to St. Louis bearable.
{\vspace{10mm} \\}
This work was supported in part by the National Science Foundation under Grant Numbers CNS-1565314 and CNS-1838271. Any opinions, findings, and conclusions or recommendations expressed in this material are my own and do not necessarily reflect the views of the National Science Foundation.
}

@@

%% Lessons learned, and future directions for research
\chapter{Conclusions on Modeling and Approximating Variability in Computer Systems} \label{ch:concs}
\input{8-conclusions}

@@

% Add your appendices here. You must leave the appendices enclosed in the appendices environment in order for the table of contents to be correct.
\begin{appendices}
\chapter{Error Bound Empirical Test Results} \label{app:error}

\input{9-error-app}

\chapter{\texttt{MQSI} Package} \label{app:mqsi}

%% -------------------------------------------------------------------
%%     Setup for including a code listing in a paper.

%% Allow TT fonts to be scaled with LaTeX macros like '\tiny'.
\newcommand{\ttscale}{s*[0.8]}
\DeclareFontFamily{OT1}{cmtt}{\hyphenchar\font -1 }
\DeclareFontShape{OT1}{cmtt}{m}{n}{
        <-9>   \ttscale cmtt8
        <9-10> \ttscale cmtt9
        <10-12>\ttscale cmtt10
        <12->  \ttscale cmtt12
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{bf}{
        <->    \ttscale cmbftt10
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{it}{
        <->    \ttscale cmitt10
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{sl}{
        <->    \ttscale cmsltt10
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{sc}{
        <->    \ttscale cmtcsc10
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{ui}
       {<->ssub*cmtt/m/it}{}
\DeclareFontShape{OT1}{cmtt}{bx}{n}
       {<->ssub*cmtt/m/n}{}
\DeclareFontShape{OT1}{cmtt}{bx}{it}
       {<->ssub*cmtt/m/it}{}
\DeclareFontShape{OT1}{cmtt}{bx}{ui}
       {<->ssub*cmtt/m/it}{}

%% Define a new color that is subtle but can be used for keywords.
\definecolor{bluegray}{rgb}{0.2, 0.3, 0.5}

\lstset{% general command to set parameter(s)
  frame=,                                  % disable the frame lines
  basicstyle=\color{black}\small\ttfamily, % whole listing style
  keywordstyle=\color{bluegray},           % language keyword color
  identifierstyle=,                        % no identifier
  commentstyle=\color{gray},               % comment color
  showstringspaces=false,                  % no special string spaces
  numbers=none                             % number lines
}
  %% numberstyle=\tiny,                       % make style of numbers tiny
  %% stepnumber=3,                            % show numbers every * lines
  %% numbersep=5pt%

%% -------------------------------------------------------------------

\lstinputlisting{MQSI/README}
\lstinputlisting[language=Fortran]{MQSI/REAL_PRECISION.f90}
\lstinputlisting[language=Fortran]{MQSI/MQSI.f90}
\lstinputlisting[language=Fortran]{MQSI/SPLINE.f90}
\lstinputlisting[language=Fortran]{MQSI/EVAL_BSPLINE.f90}

\chapter{Box Spline Subroutine} \label{app:boxspline}

\lstinputlisting[language=Fortran]{boxspline.f90}
